---
title: "PGLS analysis"
author: "Tia Harrison"
date: "`r format(Sys.Date())`"
output: html_document 
---

## Overall setup

### Global setup, load relevant packages 

```{r setup, include=FALSE}

# Global setup
knitr::opts_chunk$set(echo = TRUE)

# Packages 
library(tidyverse)
library(car)
library(lsmeans)
library(lme4)
library(MASS)
library(ape)
library(inauguration)
library(ggpubr)
library(plotrix)
library(nlme)
library(geiger)

```

### Dataset 

## Ploidy dataset and phylogeny 

The spreadsheet contains information about the symbiotic status (Fixer column) of each legume species where 1 indicates a legume that forms nodules with rhizobia (symbiotic) and where 0 indicates a plant that forms no association with rhizobia (nonsymbiotic). The diPloidyLow column indicates whether plant species are diploids (coded as 0) or polyploid (coded as 1). The diPloidyLow values were calculated from genus-level data. The Num_Introduced column indicates how many separate ranges a plant species has been introduced to as calculated in Simonsen et al (2017). 

The phylogeny used in this analysis is a version of the Zanne et al (2014) angiosperm phylogeny where the full phylogeny was pruned for the species in our dataset. 


```{r}
# Dataset 
ploidy<- read.csv("Legume_ploidy_dataset.csv", row.names=1)

# Phylogeny 
legumes1<-read.tree("Vascular_Plants_rooted.dated.tre")

```


### Clean up the dataset

Here, we create a new column called "NewPloidy" which indicates diploidy (0) and polyploidy (1) but was calculated from a combination of genus and subfamily level data. 

We also added symbiotic species to our Fixer column if we had data from Harrison et al. (2018) that suggested those species associated with a rhizobium genus or OTU. 

```{r}

# If diPloidyLow (genus-level ploidy) is available use it and if not then use disfPloidy (subfamily)
# If there is data for NumGenera or OTUs then call the associating legume species a fixer 
ploidy1<-ploidy%>% 
  rownames_to_column() %>%
  mutate(NewPloidy = ifelse(is.na(diPloidyLow), disfPloidy, diPloidyLow)) %>% 
  mutate(Fixer = ifelse(is.na(Fixer) & numOTUs >= 1, 1, Fixer)) %>%
  mutate(Fixer = ifelse(is.na(Fixer) & numGenera >=1 , 1, Fixer)) %>%
  column_to_rownames()

```


### Prune the dataset and tree 

```{r}
# Make sure the data and tree match up 
TreeOnly <- setdiff(legumes1$tip.label, rownames(ploidy1))
DataOnly <- setdiff(rownames(ploidy1), legumes1$tip.label)

# Prune the tree 
pruned <- drop.tip(legumes1, TreeOnly)
pruned$node.label <- NULL

# Get the tip names 
tip_names<-data.frame(pruned$tip.label)

# Filter the dataset for tree species 
ploidy_data<-ploidy1 %>%
  filter(rownames(ploidy1) %in% tip_names$pruned.tip.label) 

# Set columns as factors for analysis 
# Set Species as its own column
ploidy_data1 <- ploidy_data %>%
  rownames_to_column("Species") %>%
  mutate(Fixer=as.factor(Fixer), 
         numGenera=as.factor(numGenera), 
         Specialist=as.factor(Specialist), 
         diPloidyLow=as.factor(diPloidyLow), 
         NewPloidy=as.factor(NewPloidy))
```

### Data exploration 

How many species in each data type? 

```{r}
# Data summary for genus level ploidy 
ploidy_data1 %>%
  filter(!is.na(diPloidyLow) & !is.na(Num_Introduced) & !is.na(Fixer)) %>%
  group_by(diPloidyLow, Fixer) %>%
  tally()

# Average of introduced ranges 
ploidy_data1 %>%
  filter(!is.na(diPloidyLow) & !is.na(Num_Introduced) & !is.na(Fixer)) %>%
  group_by(diPloidyLow, Fixer) %>%
  summarize(mean(Num_Introduced))

# What are the two non-symbiotic polyploid species? 
ploidy_data1 %>% 
  filter(diPloidyLow==1 & Fixer ==0)
# Cassia fistula 7 human uses 
# Cassia grandis 4 human uses 

# Average number of human uses 
ploidy_data1 %>%
  filter(!is.na(Fixer) & !is.na(Human_Uses)) %>%
  group_by(Fixer) %>%
  summarize(mean(Human_Uses))

# Data summary for subfamily level ploidy data 
ploidy_data1 %>%
  filter(!is.na(NewPloidy) & !is.na(Num_Introduced) & !is.na(Fixer)) %>%
  group_by(NewPloidy, Fixer) %>%
  tally()

```

## Comparison of genus level and genus+subfamily data 
Here we looked at the species added in the genus+subfamily dataset in terms of fixer status and number of introduced ranges. 

```{r}
add_spp<- ploidy_data1 %>%
  filter(!is.na(NewPloidy) & is.na(diPloidyLow) & !is.na(Num_Introduced) & !is.na(Fixer))

add_spp %>% 
  group_by(Fixer, NewPloidy) %>% 
  tally()

add_spp %>% 
  group_by(Fixer, NewPloidy) %>%
  summarize(mean(Num_Introduced))
```


## Data analysis 

### Correlation tests 

We looked at the correlation between subfamily base chromosome number and genus base chromosome number to determine how similar our two datasets are. 

```{r}

# Spearman's rank correlation 
# Ties present so we didn't compute exact p value 
cor.test(ploidy_data1$PloidyLow, ploidy_data1$sfPloidy, method="spearman", exact=FALSE)

# Kendall tau correlation to deal with ties 
cor.test(ploidy_data1$PloidyLow, ploidy_data1$sfPloidy, method="kendall")

```


### PGLS analyses on genus-level ploidy data

We performed all analyses on our two datasets separately: genus level ploidy data and subfamily level ploidy data. We tested the hypothesis that ploidy and symbiotic status interact to impact range expansion using pgls models with the nlme package. 

```{r}
# Check that names in data and tree match  
name.check(ploidy_data1, pruned) # Says names are ok 

# Run pgls and allow lambda to vary in symbiotic status 
model_genus2<-gls(log(Num_Introduced+1) ~ diPloidyLow*Fixer + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data1, correlation=corPagel(value=0, phy=pruned, fixed=FALSE, form=~"Species"), method = "ML", na.action=na.exclude)

# Inspect model fit 
plot(model_genus2)
qqnorm(resid(model_genus2))
qqline(resid(model_genus2))

# Test for signficance 
summary(model_genus2)
Anova(model_genus2, type=3) 
# Ploidy and fixer estimates are as expected based on the raw data
# The interaction is non-significant 

# Grab means from the model 
model_genus2_results<-lsmeans(model_genus2, pairwise ~ diPloidyLow + Fixer, mode = "df.error")
model_genus2_results<- data.frame(model_genus2_results$lsmeans)


# Run pgls for specificity on rhizobia 
model_genus_spec2<-gls(log(Num_Introduced+1) ~ diPloidyLow*Specialist + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data1, correlation=corPagel(value=0.5, phy=pruned, fixed=FALSE, form=~"Species"), method = "ML", na.action=na.exclude)

# Inspect model fit 
plot(model_genus_spec2)
qqnorm(resid(model_genus_spec2))
qqline(resid(model_genus_spec2))

# Test for significance 
summary(model_genus_spec2)
Anova(model_genus_spec2, type=3)
# Ploidy estimate has the expected sign but the specialist estimate sign is opposite of what is expected 
# Interaction term is significant in this model 

# Grab the means from the model 
model_genus_spec2_results<-lsmeans(model_genus_spec2, pairwise~diPloidyLow+Specialist, mode = "df.error")
model_genus_spec2_results<- data.frame(model_genus_spec2_results$lsmeans)
model_genus_spec2_results

```

### PGLS analyses on genus-level ploidy data with simple linear model 
Run the data through a model without the phylogenetic correction to see how much the phylogenetic differences between species impacts the overall results. 

```{r}

# Run linear model on symbiotic status 
model_genus_lm<-lm(log(Num_Introduced+1) ~ diPloidyLow*Fixer + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data1)

# Inspect model fit 
plot(model_genus_lm)

# Test for signficance 
summary(model_genus_lm)
Anova(model_genus_lm, type =3) 
# The sign of the ploidy estimate is as expexted but symbiotic status is opposite to what is expected without the phylogenetic correction 
# Interaction non-significant 

# Run simple linear model for specificity on rhizobia 
model_genus_spec_lm<-lm(Num_Introduced ~ diPloidyLow*Specialist + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data1)

# Inspect model fit 
plot(model_genus_spec_lm) # QQ plot looks pretty good with raw values 

# Test for significance 
summary(model_genus_spec_lm)
Anova(model_genus_spec_lm, type=3) 
# Ploidy estimate sign is as expected, specialist sign is still opposite to what is expected
# Interaction is signficant 

```

### PGLS on the genus + subfamily level ploidy data  

Here we are testing the same hypothesis that ploidy and symbiotic status or rhizobia specificity impact range expansion. This dataset has more data points in each category of data. 

If having trouble getting the model to converge try setting different values for lambda (0, 1, or 0.5). 
If changing lambda doesn't work try this code to get convergence in the model: 
control=glsControl(opt="optim",optimMethod="Nelder-Mead") 

```{r}
# Run pgls and allow lambda to vary in symbiotic status
model_sub2<-gls(log(Num_Introduced+1) ~ NewPloidy*Fixer + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data1, correlation=corPagel(value=0, phy=pruned, fixed=FALSE, form=~"Species"), method = "ML", na.action=na.exclude)

# Inspect model fit 
plot(model_sub2)
qqnorm(resid(model_sub2))
qqline(resid(model_sub2))

# Test for significance 
summary(model_sub2)
Anova(model_sub2, type=3)
# Ploidy estimates have the expected sign, interaction is non-significant, but fixer estimate is opposite to what is expected 

# Get means from the model 
model_sub2_results<-lsmeans(model_sub2, pairwise~NewPloidy+Fixer, mode = "df.error")
model_sub2_results<-data.frame(model_sub2_results$lsmeans)


# Run pgls for specificity on rhizobia 
model_sub_spec2<-gls(log(Num_Introduced+1) ~ NewPloidy*Specialist + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data1, correlation=corPagel(value=0.5, phy=pruned, fixed=FALSE, form=~"Species"), method = "ML", na.action=na.exclude)

# Inspect model fit 
plot(model_sub_spec2)
qqnorm(resid(model_sub_spec2))
qqline(resid(model_sub_spec2))

# Test for significance 
summary(model_sub_spec2)
Anova(model_sub_spec2, type=3)
# Ploidy estimate sign is correct, specialist estimate is opposite sign as to what expected, interaction is significant 

# Grab the model means 
model_sub_spec2_results<-lsmeans(model_sub_spec2, pairwise~NewPloidy+Specialist, mode = "df.error")
model_sub_spec2_results<-data.frame(model_sub_spec2_results$lsmeans)


```


### PGLS analyses on genus+subfamily level ploidy data with simple linear model 

Run the expanded dataset through a model without the phylogenetic correction. 

```{r}

# Run linear model on symbiotic status 
sub_genus_lm<-lm(log(Num_Introduced+1) ~ NewPloidy*Fixer + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data1)

# Inspect model fit 
plot(sub_genus_lm)

# Test for signficance 
summary(sub_genus_lm)
Anova(sub_genus_lm, type=3) 
# Ploidy estimate sign is correct, Fixer sign is opposite to what is expected, and interaction is non-significant 

# Run simple linear model for specificity on rhizobia 
model_sub_spec_lm<-lm(Num_Introduced ~ NewPloidy*Specialist + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data1)

# Inspect model fit 
plot(model_sub_spec_lm) # QQ plot looks pretty good with raw values 

# Test for significance 
summary(model_sub_spec_lm)
Anova(model_sub_spec_lm, type=3) 

# Ploidy estimate sign is correct, fixer sign is opposite, significant interaction 

```

### Analyzing the non-symbiotic species 

Here we subset the data to only the non-symbiotic species in the combined genus+subfamily dataset to look for an effect of ploidy on invasion. 

```{r}
# Subset the data to non-symbiotic species 
ploidy_data_non <- ploidy_data1 %>%
  filter(Fixer == 0)
ploidy_data_non # Reduces the dataset to 68 species and two categories (diploid and polyploid)

# Run pgls and allow lambda to vary in symbiotic status
model_sub_non<-gls(log(Num_Introduced + 1) ~ NewPloidy + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data_non, correlation=corPagel(value=0, phy=pruned, fixed=FALSE, form=~"Species"), method = "ML", na.action=na.exclude)

# Inspect model fit 
plot(model_sub_non)
qqnorm(resid(model_sub_non))
qqline(resid(model_sub_non)) # This looks pretty ok with the raw data but maybe a bit better with the logged values 

# Test for significance 
summary(model_sub_non)
Anova(model_sub_non, type=3)
# Not significant but the estimate is in the direction we would expect and is getting close to significance (p = 0.0977)
```

### Analyzing the generalist species 

Here we filter the dataset for the species that are generalists and look for an impact of ploidy on range expansion. 

```{r}
# Subset the data to generalist species 
ploidy_data_gen <- ploidy_data1 %>%
  filter(Specialist == 0 & !is.na(Num_Introduced))
ploidy_data_gen # Reduced data to 43 species 

# Set the model 
model_sub_gen<-gls(Num_Introduced ~ NewPloidy + AbsLatNative + areaNativeScale + Human_Uses, data=ploidy_data_gen, correlation=corPagel(value=1, phy=pruned, fixed=FALSE, form=~"Species"), method = "ML", na.action=na.exclude)

# Inspect model fit 
plot(model_sub_gen)
qqnorm(resid(model_sub_gen))
qqline(resid(model_sub_gen)) # This looks good 

# Test for significance 
summary(model_sub_gen)
Anova(model_sub_gen, type=3)
# Overall conclusion that generalists only spread to more places around the globe when they are polyploids not diploids 
# Diploids restrict range limit severely 

```


## Plotting the results 

### Plot symbiotic status 
Plot the difference means of symbiotic and non-symbiotic ranges to see if this dataset captures results from Simonsen et al (2017). 

Our results (which is a subsetted dataset) do look like the non-symbiotic species are introduced to more ranges just like in Simonsen et al 2017. 

```{r}
# Quick plot of symbiotic status 
ggplot(ploidy_data1, aes(x=Fixer, y=Num_Introduced)) +
  geom_boxplot() + 
  theme_classic()

```

### Plot specificity on rhizobia 
Plot the difference in ranges between specialist plants and generalist plants. 

Our subsetted dataset also suggest that overall generalist plants are introduced to more ranges overall (like in Harrison et al 2018). 

```{r}
# Quick plot of specificity on rhizobia 
ploidy_data1 %>% 
  filter(!is.na(Specialist)) %>%
  ggplot(aes(x=Specialist, y=Num_Introduced)) +
  geom_boxplot() + 
  theme_classic()

```

### Interaction of ploidy and symbiosis in genus level data 

Plot the means and standard error calculated from the raw data. 

```{r}

# Prep the data for plotting 
# Calculate raw means and standard error for plot 
ploidy_data_sum<- ploidy_data1 %>%
  filter(!is.na(Fixer) & !is.na(diPloidyLow)) %>%
  dplyr::group_by(diPloidyLow, Fixer) %>%
  dplyr::summarise(mean=mean(Num_Introduced), se=std.error(Num_Introduced))

# Get colours 
ploidy_colours<-inauguration("inauguration_2021")

# Reaction norm plot 
genus_interaction<-ggplot(ploidy_data_sum, aes(x=Fixer, y=mean, fill=diPloidyLow)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1, position=position_dodge(0.2)) +
  geom_line(aes(group=diPloidyLow), position=position_dodge(0.2)) +
  geom_point(pch =21, size=4, position=position_dodge(0.2)) + 
  (ylab("Introduced ranges (no.)")) +
  (xlab("Symbiotic state")) +
  scale_fill_manual(name="", labels=c("Diploid", "Polyploid"), values=ploidy_colours[c(6,4)]) +
  scale_x_discrete(labels=c("1" = "Symbiotic", "0" = "Non-symbiotic"), limits=c("1", "0")) + 
  theme_classic() +
  theme(axis.title.x = element_text(size=13), 
        axis.title.y = element_text(size=13), 
        axis.text.y= element_text(size=12), 
        axis.text.x= element_text(size=12),
        legend.title= element_text(size=12), 
        legend.text=element_text(size=12), 
        legend.position="top")

# Prep data for specificity on rhizobia 
ploidy_data_sum2<- ploidy_data1 %>%
  filter(!is.na(Specialist) & !is.na(diPloidyLow)) %>%
  dplyr::group_by(diPloidyLow, Specialist) %>%
  dplyr::summarise(mean=mean(Num_Introduced), se=std.error(Num_Introduced))

# Reaction norm for specificity 
genus_spec<-ggplot(ploidy_data_sum2, aes(x=Specialist, y=mean, fill=diPloidyLow)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1, position=position_dodge(0.2)) +
  geom_line(aes(group=diPloidyLow), position=position_dodge(0.2)) +
  geom_point(pch =21, size=4, position=position_dodge(0.2)) + 
  (ylab("Introduced ranges (no.)")) +
  (xlab("Specificity on rhizobia")) +
  scale_fill_manual(name="", labels=c("Diploid", "Polyploid"), values=ploidy_colours[c(6,4)]) +
  scale_x_discrete(labels=c("1" = "Specialist", "0" = "Generalist"), limits=c("1", "0")) + 
  theme_classic() +
  theme(axis.title.x = element_text(size=13), 
        axis.title.y = element_text(size=13), 
        axis.text.y= element_text(size=12), 
        axis.text.x= element_text(size=12),
        legend.title= element_text(size=12), 
        legend.text=element_text(size=12), 
        legend.position="top")

# Combine the two plots 
Figure_genus <- ggarrange(genus_interaction, genus_spec,
                    labels=c("a", "b"), 
                    ncol = 2, nrow=1,
                    common.legend=TRUE, 
                    legend = "top")
Figure_genus

```

### PGLS interaction of ploidy and symbiosis in genus level data

The means and standard error values plotted here are taken from the pgls model (nlme model) from our analyses. 

```{r}

# Prep the model output 
model_genus2_results <- model_genus2_results %>%
  mutate(diPloidyLow=as.factor(diPloidyLow)) %>%
  mutate(Fixer=as.factor(Fixer))

# Plot the model means 
genus_interaction2<-ggplot(model_genus2_results, aes(x=Fixer, y=lsmean, fill=diPloidyLow)) + 
  geom_errorbar(aes(ymin=lsmean-SE, ymax=lsmean+SE), width=0.1, position=position_dodge(0.2)) +
  geom_line(aes(group=diPloidyLow), position=position_dodge(0.2)) +
  geom_point(pch =21, size=4, position=position_dodge(0.2)) + 
  (ylab("Introduced ranges (log)")) +
  (xlab("Symbiotic state")) +
  scale_fill_manual(name="", labels=c("Diploid", "Polyploid"), values=ploidy_colours[c(6,4)]) +
  scale_x_discrete(labels=c("1" = "Symbiotic", "0" = "Non-symbiotic"), limits=c("1", "0")) + 
  theme_classic() +
  theme(axis.title.x = element_text(size=13), 
        axis.title.y = element_text(size=13), 
        axis.text.y= element_text(size=12), 
        axis.text.x= element_text(size=12),
        legend.title= element_text(size=12), 
        legend.text=element_text(size=12), 
        legend.position="top")

# Prep the model output 
model_genus_spec2_results <- model_genus_spec2_results %>%
  mutate(diPloidyLow=as.factor(diPloidyLow)) %>%
  mutate(Specialist=as.factor(Specialist))

# Plot the model means 
genus_spec2<-ggplot(model_genus_spec2_results, aes(x=Specialist, y=lsmean, fill=diPloidyLow)) + 
  geom_errorbar(aes(ymin=lsmean-SE, ymax=lsmean+SE), width=0.1, position=position_dodge(0.2)) +
  geom_line(aes(group=diPloidyLow), position=position_dodge(0.2)) +
  geom_point(pch =21, size=4, position=position_dodge(0.2)) + 
  (ylab("Introduced ranges (log)")) +
  (xlab("Specificity on rhizobia")) +
  scale_fill_manual(name="", labels=c("Diploid", "Polyploid"), values=ploidy_colours[c(6,4)]) +
  scale_x_discrete(labels=c("1" = "Specialist", "0" = "Generalist"), limits=c("1", "0")) + 
  theme_classic() +
  theme(axis.title.x = element_text(size=13), 
        axis.title.y = element_text(size=13), 
        axis.text.y= element_text(size=12), 
        axis.text.x= element_text(size=12),
        legend.title= element_text(size=12), 
        legend.text=element_text(size=12), 
        legend.position="top")

# Combine the two plots 
Figure_genus2 <- ggarrange(genus_interaction2, genus_spec2,
                    labels=c("a", "b"), 
                    ncol = 2, nrow=1,
                    common.legend=TRUE, 
                    legend = "top")
Figure_genus2

```


### Interaction of ploidy and symbiosis in genus+subfamily level data 

Here we plot the means and standard error calculated from the raw data in the expanded dataset (more data points in each category). 

```{r}
# Prep the data for plotting 
ploidy_sub_sum<- ploidy_data1 %>%
  filter(!is.na(Fixer) & !is.na(NewPloidy)) %>%
  dplyr::group_by(NewPloidy, Fixer) %>%
  dplyr::summarise(mean=mean(Num_Introduced), se=std.error(Num_Introduced))

# Reaction norm plot 
sub_interaction<-ggplot(ploidy_sub_sum, aes(x=Fixer, y=mean, fill=NewPloidy)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1, position=position_dodge(0.2)) +
  geom_line(aes(group=NewPloidy), position=position_dodge(0.2)) +
  geom_point(pch =21, size=4, position=position_dodge(0.2)) + 
  (ylab("Introduced ranges (no.)")) +
  (xlab("Symbiotic state")) +
  scale_fill_manual(name="", labels=c("Diploid", "Polyploid"), values=ploidy_colours[c(6,4)]) +
  scale_x_discrete(labels=c("1" = "Symbiotic", "0" = "Non-symbiotic"), limits=c("1", "0")) + 
  theme_classic() +
  theme(axis.title.x = element_text(size=13), 
        axis.title.y = element_text(size=13), 
        axis.text.y= element_text(size=12), 
        axis.text.x= element_text(size=12),
        legend.title= element_text(size=12), 
        legend.text=element_text(size=12), 
        legend.position="top")

# Prep data for specificity on rhizobia 
ploidy_sub_spec<- ploidy_data1 %>%
  filter(!is.na(Specialist) & !is.na(NewPloidy)) %>%
  dplyr::group_by(NewPloidy, Specialist) %>%
  dplyr::summarise(mean=mean(Num_Introduced), se=std.error(Num_Introduced))

# Reaction norm for specificity 
sub_spec<-ggplot(ploidy_sub_spec, aes(x=Specialist, y=mean, fill=NewPloidy)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1, position=position_dodge(0.2)) +
  geom_line(aes(group=NewPloidy), position=position_dodge(0.2)) +
  geom_point(pch =21, size=4, position=position_dodge(0.2)) + 
  (ylab("Introduced ranges (no.)")) +
  (xlab("Specificity on rhizobia")) +
  scale_fill_manual(name="", labels=c("Diploid", "Polyploid"), values=ploidy_colours[c(6,4)]) +
  scale_x_discrete(labels=c("1" = "Specialist", "0" = "Generalist"), limits=c("1", "0")) + 
  theme_classic() +
  theme(axis.title.x = element_text(size=13), 
        axis.title.y = element_text(size=13), 
        axis.text.y= element_text(size=12), 
        axis.text.x= element_text(size=12),
        legend.title= element_text(size=12), 
        legend.text=element_text(size=12), 
        legend.position="top")

# Combine the two plots 
Figure_sub <- ggarrange(sub_interaction, sub_spec,
                    labels=c("a", "b"), 
                    ncol = 2, nrow=1,
                    common.legend=TRUE, 
                    legend = "top")
Figure_sub
```

### PGLS interaction of ploidy and symbiosis in genus+subfamily level data

The means and standard deviation plotted here are output from the pgls models in our analyses. 

```{r}

# Prep the output from the model  
model_sub2_results <- model_sub2_results %>%
  mutate(NewPloidy=as.factor(NewPloidy)) %>%
  mutate(Fixer=as.factor(Fixer))
model_sub2_results

# Plot the means 
sub_interaction2<-ggplot(model_sub2_results, aes(x=Fixer, y=lsmean, fill=NewPloidy)) + 
  geom_errorbar(aes(ymin=lsmean-SE, ymax=lsmean+SE), width=0.1, position=position_dodge(0.2)) +
  geom_line(aes(group=NewPloidy), position=position_dodge(0.2)) +
  geom_point(pch =21, size=4, position=position_dodge(0.2)) + 
  (ylab("Introduced ranges (log)")) +
  (xlab("Symbiotic state")) +
  scale_fill_manual(name="", labels=c("Diploid", "Polyploid"), values=ploidy_colours[c(6,4)]) +
  scale_x_discrete(labels=c("1" = "Symbiotic", "0" = "Non-symbiotic"), limits=c("1", "0")) + 
  theme_classic() +
  theme(axis.title.x = element_text(size=13), 
        axis.title.y = element_text(size=13), 
        axis.text.y= element_text(size=12), 
        axis.text.x= element_text(size=12),
        legend.title= element_text(size=12), 
        legend.text=element_text(size=12), 
        legend.position="top")

# Prep the output from the model  
model_sub_spec2_results <- model_sub_spec2_results %>%
  mutate(NewPloidy=as.factor(NewPloidy)) %>%
  mutate(Specialist=as.factor(Specialist))

# Plot the model means 
sub_spec2<-ggplot(model_sub_spec2_results, aes(x=Specialist, y=lsmean, fill=NewPloidy)) + 
  geom_errorbar(aes(ymin=lsmean-SE, ymax=lsmean+SE), width=0.1, position=position_dodge(0.2)) +
  geom_line(aes(group=NewPloidy), position=position_dodge(0.2)) +
  geom_point(pch =21, size=4, position=position_dodge(0.2)) + 
  (ylab("Introduced ranges (log)")) +
  (xlab("Specificity on rhizobia")) +
  scale_fill_manual(name="", labels=c("Diploid", "Polyploid"), values=ploidy_colours[c(6,4)]) +
  scale_x_discrete(labels=c("1" = "Specialist", "0" = "Generalist"), limits=c("1", "0")) + 
  theme_classic() +
  theme(axis.title.x = element_text(size=13), 
        axis.title.y = element_text(size=13), 
        axis.text.y= element_text(size=12), 
        axis.text.x= element_text(size=12),
        legend.title= element_text(size=12), 
        legend.text=element_text(size=12), 
        legend.position="top")

# Combine the plots 
Figure_sub2 <- ggarrange(sub_interaction2, sub_spec2,
                    labels=c("a", "b"), 
                    ncol = 2, nrow=1,
                    common.legend=TRUE, 
                    legend = "top")
Figure_sub2

```





