---
title: "State transitions analysis"
author: "Tia Harrison"
date: "`r format(Sys.Date())`"
output: github_document 
editor_options: 
  chunk_output_type: console
---

## Overall setup

### Setup 

```{r setup, include=FALSE}

# Global setup
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

```


### Load the packages for analysis 

```{r packages}

# Packages 
library(tidyverse)
library(corHMM)
library(ape)
library(expm)
library(hablar)
library(dplyr)
library(phytools)
library(devtools)
library(inauguration)
library(diagram)
```


### Ploidy dataset and phylogeny 

The datasheet contains information about the symbiotic status (Fixer column) of the legume species where 1 indicates a legume that forms nodules with rhizobia (aka a mutualist) and where 0 indicates a plant that forms no association with rhizobia (nonmutualist). The diPloidyLow column indicates whether plant species are diploids (coded as 0) or polyploid (coded as 1). The diPloidyLow values were calculated from genus-level data. The NewPloidy column also indicates diploidy (0) and polyploidy (1) but was calculated from a combination of genus and subfamily level data.  


```{r data}
# Dataset 
ploidy<- read.csv("Legume_ploidy_dataset.csv", row.names=1)

# Phylogeny 
legumes1<-read.tree("Vascular_Plants_rooted.dated.tre")

# Clean up the data 
ploidy1<-ploidy%>%
  rownames_to_column() %>%
  mutate(NewPloidy = ifelse(is.na(diPloidyLow), disfPloidy_corrected, diPloidyLow)) %>%
  mutate(Fixer = ifelse(is.na(Fixer) & numOTUs >= 1, 1, Fixer)) %>%
  mutate(Fixer = ifelse(is.na(Fixer) & numGenera >=1 , 1, Fixer)) %>%
  column_to_rownames()

```


### Pruning for the genus + subfamily level ploidy data  

We pruned the Zanne et al (2014) angiosperm phylogeny for legume species in our dataset.


```{r prune}
# Remove all NAs from the data for the relevant columns 
ploidy_sub <- ploidy1 %>% 
  filter(!is.na(Fixer), !is.na(NewPloidy))

# Make sure the data and tree match up 
TreeOnly_sub <- setdiff(legumes1$tip.label, rownames(ploidy_sub))
DataOnly_sub <- setdiff(rownames(ploidy_sub), legumes1$tip.label)

# Prune the tree 
sub_tree <- drop.tip(legumes1, TreeOnly_sub)
sub_tree$node.label <- NULL

# Get the tip names 
tip_names_sub<-data.frame(sub_tree$tip.label)

# Filter the dataset for tree species 
ploidy_sub2<-ploidy_sub %>%
  rownames_to_column("Species") %>%
  filter(Species %in% tip_names_sub$sub_tree.tip.label) %>%
  dplyr::select(Species, Fixer, NewPloidy)

```


## Estimating state transitions on subfamily and genus level ploidy data 

### Simple model 

We ran a simple model first with no hidden states. Dual transitions are not allowed in the model, only one state transition at a time. We allowed polyploids to revert back to diploids in the model. 

Results are interpreted as transitions from ROW to COLUMN 
Traits:
1 = nonmutualist diploid 
2 = nonmutualist polyploid 
3 = mutualist diploid 
4 = mutualist polyploid 

```{r simple model}
# Set seed for reproducibility 135
set.seed(5348)

# Simple model with two traits and no hidden states 
MK_sub_simple<- corHMM(phy=sub_tree, data=ploidy_sub2, rate.cat=1)

# Look at the results of the model 
MK_sub_simple
plotMKmodel(MK_sub_simple)

```

### Hidden states model 

We included rate.cat=2 to specify hidden states in the model. This model with hidden states performed better than the model with no hidden states and so it is reported in the manuscript. 

Results are interpreted as transitions from ROW to COLUMN 
Traits:
1 = nonmutualist diploid 
2 = nonmutualist polyploid 
3 = mutualist diploid 
4 = mutualist polyploid 

States: 
R1 = rate category 1
R2 = rate category 2 

```{r hidden model}

# Two traits and hidden states model 
MK_sub_hidden<- corHMM(phy=sub_tree, data=ploidy_sub2, rate.cat=2)

# Look at the results of the model 
MK_sub_hidden

# Make two colours for gradient 
state_colours1<- inauguration("inauguration_2021")
colour_grad<-state_colours1[c(1,6)]

# Look at the rates 
plotMKmodel(MK_sub_hidden, color=colour_grad, vertex.scale= 0.5, arrow.scale= 0.8, display = "square")

# Prep data for plotting diagram 
rate_data<-as.matrix(MK_sub_hidden$solution) # Make a matrix 
rate_dataR1<-rate_data[1:4, 1:4] # Select the R1 pieces of the matrix 
rate_dataR1<-round(rate_dataR1, 5) # Round the values in the matrix
rate_dataR1<-t(rate_dataR1) # Transpose the matrix so the direction is correct

rate_dataR2<-rate_data[5:8, 5:8] # Select the R2 pieces of the matrix 
rate_dataR2<-round(rate_dataR2, 5) # Round values 
rate_dataR2<-t(rate_dataR2) # Transpose matrix 

# Manipulate matrix for arrow line width plotting 
R1_thick<-1+rate_dataR1 
R2_thick<-1+rate_dataR2

# Set up plot save 
pdf("Figures/Figure_R1.pdf") # To save the figure 

# Make the plot 
R1_plot<-plotmat(rate_dataR1, pos=c(1,2,1),
              name = c(" non-symbiotic \ndiploid", " non-symbiotic \npolyploid", " symbiotic \ndiploid",  " symbiotic \npolyploid"),
              box.lwd = 1.5, cex.txt = 0.75, curve=0.04, box.col=c("white", "white", "lightgrey", "lightgrey"), shadow.size=0, arr.lwd=R1_thick, arr.type="triangle", 
              box.size = 0.1, box.type = "square", box.prop = 0.5,
              main = "Rate Category 1")

dev.off()

# Set up plot save 
pdf("Figures/Figure_R2.pdf")

R2_plot<-plotmat(rate_dataR2, pos=c(1,2,1),
              name = c(" non-symbiotic \ndiploid", " non-symbiotic \npolyploid", " symbiotic \ndiploid",  " symbiotic \npolyploid"),
              box.lwd = 1.5, cex.txt = 0.75, curve=0.04, box.col=c("white", "white", "lightgrey", "lightgrey"), shadow.size=0, arr.lwd=R2_thick, arr.type="triangle", 
              box.size = 0.1, box.type = "square", box.prop = 0.5,
              main = "Rate Category 2")

dev.off()


```


## Visualization of transitions

### Prep data for visualization 

Pick 8 colours for plotting the different transitions on the tree and re-label to the corresponding states for ploidy and symbiosis. 

```{r prep}

# Load colour palettes 
state_colours2<- inauguration("bernie_mittens")

# Get the colour codes 
color1<-state_colours1[c(2,5,3)]
color2<-state_colours2[c(2)]

# Darken and lighten the colours 
dark1=colorRampPalette(c("#749dae", "black"))(10)
lite1=colorRampPalette(c("#749dae", "white"))(10)
dark2=colorRampPalette(c("#cd3341", "black"))(10)
lite2=colorRampPalette(c("#cd3341", "white"))(10)
dark3=colorRampPalette(c("#50506D", "black"))(10)
lite3=colorRampPalette(c("#50506D", "white"))(10)
dark4=colorRampPalette(c("#f3c483", "black"))(10)
lite4=colorRampPalette(c("#f3c483", "white"))(10)

# Combine colours for for plotting 
state_colours8<-setNames(c("#5A7A87", "#9F2732", "#3E3E54", "#BD9865", "#A2BDC9", "#DD7780", "#8A8A9D", "#F7D7AC"),
                         c("(0_0,R1)", "(0_1,R1)", "(1_0,R1)", "(1_1,R1)", "(0_0,R2)", "(0_1,R2)", "(1_0,R2)", "(1_1,R2)")) 

# Colours and labels for the legend
legend_colours<-setNames(c("#5A7A87", "#9F2732", "#3E3E54", "#BD9865", "#A2BDC9", "#DD7780", "#8A8A9D", "#F7D7AC"),
                         c("non-symbiotic, diploid, R1", "non-symbiotic, polyploid, R1", "symbiotic, diploid, R1", "symbiotic, polyploid, R1", "non-symbiotic, diploid, R2", "non-symbiotic, polyploid, R2", "symbiotic, diploid, R2", "symbiotic, polyploid, R2")) 

```

### Genus/Subfamily level ploidy data 

We plotted the transition rates determined from the hidden state model which was the best fitting model, for the largest dataset (genus + subfamily level ploidy). 

```{r plot}

# Extract the data from the model 
phy_sub_hidden = MK_sub_hidden$phy
data_sub_hidden = MK_sub_hidden$data
model_sub_hidden = MK_sub_hidden$solution
model_sub_hidden[is.na(model_sub_hidden)] <- 0
diag(model_sub_hidden) <- -rowSums(model_sub_hidden)

# Run to get simmap 
simmap_sub_hidden <- makeSimmap(tree = phy_sub_hidden, data = data_sub_hidden, model = model_sub_hidden, rate.cat = 2)

# Import phytools plotSimmap for plotting
phytools::plotSimmap(simmap_sub_hidden[[1]], colors=state_colours8, fsize = 0.15, offset=3, type="fan")

# Figure out the subfamilies 
unique(ploidy$subfamily)
# Papilionoideae Mimosoid Caesalpinioideae Dialioideae Cercidoideae Detarioideae

# Save the subfamily data 
subfam<-ploidy1 %>%
  select(subfamily)

# Plot the tree with subfamily labels to get an idea of where to put labels 
library(ggtree)
circ <- ggtree(simmap_sub_hidden[[1]], layout = "circular")
p1 <- gheatmap(circ, subfam, offset=.8, width=.2,
               colnames_angle=95, colnames_offset_y = .25) +
    scale_fill_viridis_d(option="D", name="discrete\nvalue")
p1

# Get the node labels 
phytools::plotSimmap(simmap_sub_hidden[[1]], colors=state_colours8, fsize = 0.15, offset=3, type="fan")
nodelabels(cex=0.18)
# Put the subfamily labels on the black and white figure below 

# Set up plot save the phylogeny with the full information 
pdf("Figures/Figure_transition_tree_full.pdf", width =9, height =6.6)

# Run to get simmap 
simmap_sub_hidden <- makeSimmap(tree = phy_sub_hidden, data = data_sub_hidden, model = model_sub_hidden, rate.cat = 2)

# Plot the tree with the proper labels 
phytools::plotSimmap(simmap_sub_hidden[[1]], colors=state_colours8, fsize = 0.15, offset=3, type="fan")
arc.cladelabels(text="Papilionoideae",node=1043,ln.offset=1.15,lab.offset=1.2, cex=0.6, col="grey", mark.node=FALSE) 
# arc.cladelabels(text="Mimosoid",node=912,ln.offset=1.1,lab.offset=1.15) # Give the clade labels based on internal node labels rather than tip labels (specifying tips will be too )
arc.cladelabels(text="Cercidoideae",node=849,ln.offset=1.15,lab.offset=1.2, orientation="horizontal", cex=0.6, mark.node=FALSE)
arc.cladelabels(text="Caesalpinioideae",node=875,ln.offset=1.15,lab.offset=1.2, cex=0.6, mark.node=FALSE)
arc.cladelabels(text="Detarioideae",node=865,ln.offset=1.15,lab.offset=1.2, cex=0.6, orientation="horizontal", col="grey", mark.node=FALSE)

dev.off()

# Set up plot to get the legend

phytools::plotSimmap(simmap_sub_hidden[[1]], colors=state_colours8, fsize = 0.01, offset=3, type="fan")
phytools::add.simmap.legend(colors=legend_colours, fsize=0.7, prompt=FALSE, x=-120, y=70, vertical=TRUE)

```

Split up the figure into multiple panels with different colours. One panel that just shows the rate classes, one panel that just shows the 4 categories, and one panel that has all the information (rates and 4 categories). 

```{r plot subfamilies}
# Colours for two rate classes 
state_colours_rate<-setNames(c("black", "black", "black", "black", "grey", "grey", "grey", "grey"),
                         c("(0_0,R1)", "(0_1,R1)", "(1_0,R1)", "(1_1,R1)", "(0_0,R2)", "(0_1,R2)", "(1_0,R2)", "(1_1,R2)")) 

# Set up plot save 
pdf("Figures/Figure_transition_rates.pdf")

# Run to get simmap 
simmap_sub_hidden_R <- makeSimmap(tree = phy_sub_hidden, data = data_sub_hidden, model = model_sub_hidden, rate.cat = 2)

# Import phytools plotSimmap for plotting
# Put the subfamily labels on this one 
phytools::plotSimmap(simmap_sub_hidden_R[[1]], colors=state_colours_rate, fsize = 0.15, offset=3, type="fan")

dev.off()


# Legend for the two rate classes 

phytools::plotSimmap(simmap_sub_hidden[[1]], colors=state_colours_rate, fsize = 0.01, offset=3, type="fan")
phytools::add.simmap.legend(colors=legend_colours, fsize=0.7, prompt=FALSE, x=-120, y=70, vertical=TRUE)


# Set up plot save 
pdf("Figures/Figure_transition_4.pdf")

# Now the tree with the four classes 
state_colours4<-setNames(c("#5A7A87", "#9F2732", "#3E3E54", "#BD9865", "#5A7A87", "#9F2732", "#3E3E54", "#BD9865"),
                         c("(0_0,R1)", "(0_1,R1)", "(1_0,R1)", "(1_1,R1)", "(0_0,R2)", "(0_1,R2)", "(1_0,R2)", "(1_1,R2)")) 

# Run to get simmap 
simmap_sub_hidden_4 <- makeSimmap(tree = phy_sub_hidden, data = data_sub_hidden, model = model_sub_hidden, rate.cat = 2)

# Import phytools plotSimmap for plotting
phytools::plotSimmap(simmap_sub_hidden_4[[1]], colors=state_colours4, fsize = 0.15, offset=3, type="fan")

dev.off()

```



